---
title: "Using the Boruta Package to Determine Feature Relevance"
author: "Jim Thompson"
date: "February 13, 2016"
output: html_document
---

This report determines what features may be relevant to modeling the **target**
attribute for the BNP Paribas Cardif competition.  This analysis is based on
the [**Boruta**](https://cran.r-project.org/web/packages/Boruta/index.html) package.
In addition we demonstrate the [**caret**](https://cran.r-project.org/web/packages/caret/index.html) package functions 
to create data partitions and impute missing values.

First, we load required libraries.
```{r warning=FALSE,message=FALSE}
# required libraries
library(caret)
library(Boruta)
library(dplyr)
```

Read in the training data set, determine data types and perform simple pre-processing 
for numeric data.
```{r}
DATA.DIR <- "../data"

# retrive sample data for analysis
train <- read.csv(paste0(DATA.DIR,"/train.csv"),stringsAsFactors = FALSE)

###
# select random sample for analysis using caret createDataPartition() function
###
set.seed(123)
idx <- createDataPartition(train$target,p=0.025,list=FALSE)
sample.df <- train[idx,]

###
# segregate numeric vs character data types
###
# get names of the explanatory variables
explanatory.attributes <- setdiff(names(sample.df),c("ID","target"))

# determine data type for each explanatory variable
data.classes <- sapply(explanatory.attributes,function(x){class(sample.df[,x])})

# segregate explanatory variables by data type, eg. character, numeric, integer
unique.classes <- unique(data.classes)
attr.by.data.types <- lapply(unique.classes,function(x){names(data.classes[data.classes==x])})
names(attr.by.data.types) <- unique.classes
comment(attr.by.data.types) <- "list that categorize training data types"


# for numeric attributes use caret preProcess() and predict() functions to impute missing values
pp <- preProcess(sample.df[c(attr.by.data.types$numeric,attr.by.data.types$integer)],
                 method=c("medianImpute"))
pp.sample.df <- predict(pp,sample.df[c(attr.by.data.types$numeric,attr.by.data.types$integer)])

# combine numeric data with character data
df <- cbind(pp.sample.df,sample.df[attr.by.data.types$character])

```


Analyze sample data with Boruta package to determine relevance of the explanatory
variables.
```{r}
set.seed(13)
bor.results <- Boruta(df,factor(sample.df$target),
                   maxRuns=51,
                   doTrace=2)
```


```{r echo=FALSE}
cat("\nsummary of Boruta run:\n")
print(bor.results)

cat("\n\nAttributes determined to be relevant:\n")
getSelectedAttributes(bor.results)
```

Plot of variable importance
```{r echo=FALSE,fig.width=9,fig.height=7}
plot(bor.results)
```

Detailed results for each explanatory results
```{r, echo=FALSE}
cat("\n\nAttribute importance details:\n")
arrange(attStats(bor.results),desc(medianImp))
```


